Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-07-04 paper_jasa.tex
comments updated

This proves the statement above. 
%We further see:
%\[
%E\left[ (1 - p_D)^{m-1}\right] =  \sum_{k=0}^{m-1} {m-1 \choose k} (-1)^k E[p_D^k] = 1-(m-1)E[p_D] + O(E[p_D^2]).
%\]
This implies that as $p_D$, the $p$-value, becomes smaller, the probability that the observer identifies the data plot goes up and \blue{(How can we say that from this? The term  $E\left[ (1 - p_D)^{m-1}\right]$ is definitely not a function of $p_D$ since $p_D$ should be eliminated after the expectation. Instead, its a function of $m$ only ! Is this interesting that under the minimum $p$-value assumption the power is a function of lineup size $m$ only?)},
conversely, as $m$, the number of choices given in the lineup, goes up the probability to pick the data plot  goes down. 
}
%
%
%
%Let us think of a lineup as a head-to-head comparison of the test statistic and $m-1$ null plots. 
%%Let the data plot have a $p$-value of $p_D$, and the null plots $p$ values of $p_{0, i}$ with $1 \le i \le m-1$.
%We know that for each comparison the probability that the data, on which the plot is based, has the smaller $p$ value is 
%\begin{eqnarray*}
%P(p_D < p_{0,i}) &=& 1 - P(p_{i,0} \le p_D) = 1- \int_0^1  P(p_{i,0} \le p_D \mid p_D=t) f_{p_D}(t) dt =  \\
%&=& 1 - \int_0^1 F_{p_{0,i}}(t) f_{p_D}(t) dt = 1 - \int_0^1 t f_{p_D}(t) dt = 1 - E[p_D],
%\end{eqnarray*}
%which, in particular, is independent of $p_{0,i}$ for all $i$.
%
%Let us make the assumption that an observer is able to identify  the chart corresponding to the data with the smallest $p$-value. Further we will assume that all observers  have the same ability in identifying the data plot.
%
%With that, we define $Z$  as the number of null plots in a lineup, for which the $p$-value $p_{0,i}$ is smaller than $p_D$.  
%
%Then $Z \sim B_{m-1, E[p_D]}$, and the probability that an observer will pick the data plot in a given lineup is 
%\[
%P(Z=0) = \left(1 - E[p_D] \right)^{m-1} = P(p_D \le p_0), \ \ \ \text{ where } p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}.
%\]
\end{proof}

>>>>

This proves the statement above. 
%We further see:
%\[
%E\left[ (1 - p_D)^{m-1}\right] =  \sum_{k=0}^{m-1} {m-1 \choose k} (-1)^k E[p_D^k] = 1-(m-1)E[p_D] + O(E[p_D^2]).
%\]
This implies that as $p_D$, the $p$-value, becomes smaller, the probability that the observer identifies the data plot goes up and \blue{(How can we say that from this? The term  $E\left[ (1 - p_D)^{m-1}\right]$ is definitely not a function of $p_D$ since $p_D$ should be eliminated after the expectation. Instead, its a function of $m$ and the distribution of $p_D$ ! If the density of $p$-value is very right skewed, the expectation term would be large. The distribution function would be highly right skewed when there would be strong signal in the data plot.Thats where the distribution function of $p$-value under alternative comes in. Should we add that in Equation \ref{dist_p}? )},
conversely, as $m$, the number of choices given in the lineup, goes up the probability to pick the data plot  goes down. 
}
%
%
%
%Let us think of a lineup as a head-to-head comparison of the test statistic and $m-1$ null plots. 
%%Let the data plot have a $p$-value of $p_D$, and the null plots $p$ values of $p_{0, i}$ with $1 \le i \le m-1$.
%We know that for each comparison the probability that the data, on which the plot is based, has the smaller $p$ value is 
%\begin{eqnarray*}
%P(p_D < p_{0,i}) &=& 1 - P(p_{i,0} \le p_D) = 1- \int_0^1  P(p_{i,0} \le p_D \mid p_D=t) f_{p_D}(t) dt =  \\
%&=& 1 - \int_0^1 F_{p_{0,i}}(t) f_{p_D}(t) dt = 1 - \int_0^1 t f_{p_D}(t) dt = 1 - E[p_D],
%\end{eqnarray*}
%which, in particular, is independent of $p_{0,i}$ for all $i$.
%
%Let us make the assumption that an observer is able to identify  the chart corresponding to the data with the smallest $p$-value. Further we will assume that all observers  have the same ability in identifying the data plot.
%
%With that, we define $Z$  as the number of null plots in a lineup, for which the $p$-value $p_{0,i}$ is smaller than $p_D$.  
%
%Then $Z \sim B_{m-1, E[p_D]}$, and the probability that an observer will pick the data plot in a given lineup is 
%\[
%P(Z=0) = \left(1 - E[p_D] \right)^{m-1} = P(p_D \le p_0), \ \ \ \text{ where } p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}.
%\]
\end{proof}
